\chapter{CFAST Developers}

The Consolidated Model of Fire and Smoke Transport (CFAST) and Smokeview are the products of a collaborative effort led by the National Institute of Standards and Technology (NIST). Its developers and contributors are listed below.

\vspace{0.3in}

\begin{flushleft}

Principal Developers of CFAST  \\ [0.2in]

Richard Peacock, NIST \\
Glenn Forney, NIST \\
Paul Reneke, NIST \\
Kevin McGrattan, NIST \\
Walter Jones \\ [0.3in]

Principal Developer of Smokeview  \\ [0.2in]

Glenn Forney, NIST \\ [0.3in]

Contributors \\ [0.2in]

Kristopher Overholt, Continuum Analytics, Austin, Texas, USA \\ [0.3in]
Jason Floyd, Fire Safety Research Institute, UL Research Institutes, Columbia, Maryland, USA \\

\end{flushleft}


\chapter{About the Developers}

\begin{description}

\item[Richard Peacock] is a chemical engineering in the Fire Research Division of NIST. He received a bachelor of science degree from the Clark School of Engineering of the University of Maryland in 1973. He joined the NIST staff in 1974 (then the National Bureau of Standards) and has worked on real-scale testing and the development and validation of fire models, most notably CFAST.

\item[Glenn Forney] is a computer scientist in the Fire Research Division of NIST.  He received a bachelor of science degree in mathematics from Salisbury State College and a master of science and a doctorate in mathematics from Clemson University.  He joined NIST in 1986 (then the National Bureau of Standards) and has since worked on developing tools that provide a better understanding of fire phenomena, most notably Smokeview, an advanced scientific software tool for visualizing Fire Dynamics Simulation data.

\item[Paul Reneke] is a computer scientist in the Fire Research Division of NIST.  He received a bachelor of science degree in mathematical sciences from Clemson Univerity and a master of science degree in applied mathematics from The Johns Hopkins University. He joined NIST in 1990. He has worked on the development of user interfaces, graphics and improved numerics in fire models, notably CFAST. His research interests include sensitivity analysis and validation of fire models.

\item[Kevin McGrattan] is a mathematician in the Fire Research Division of NIST. He received a bachelor of science degree from the School of Engineering and Applied Science of Columbia University in 1987 and a doctorate at the Courant Institute of New York University in 1991. He joined the NIST staff in 1992 and has since worked on the development and validation of fire models, most notably the Fire Dynamics Simulator.

\item[Walter Jones] was a physicist at NIST (now retired). He received a bachelor of arts degree in physics from Oberlin College and a doctorate in physics from the University of Maryland. He was the original developer of the CFAST model. In addition to the development of fire models, he has worked on smart fire alarms and smoke control for naval vessels.

\end{description}


